Long-term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-08-25 article/article.Rmd
Commit 2024-08-25 17:12

<!--
We studied the enhancement of the global horizontal irradiance induced by clouds over
Thessaloniki for the period 1993 -- 2023 using data recorded every one minute.  We
identified the cloud enhancement events (CE) by creating an appropriate clear sky
irradiance reference with the use of a radiative transfer model and aerosol optical
depth data from a collocated Cimel sun photometer and a Brewer spectrophotometer. We
found a trend in CE events of
$`r round(yearlyCE[var == "GLB_diff.N", slope], 1)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.N", slope.sd], 1)`\,\text{cases}/\text{year}$,
and a trend in the CE events irradiation of 
$`r round(yearlyCE[var == "GLB_diff.sum", slope], 1)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.sum", slope.sd], 1)`\,\text{kJ}/\text{year}$.
The peak of the CE events was observed during May and June.  An analysis of the total
duration of CE events, showed that durations longer than 5 minutes are very rare,
with exceptions lasting over an hour and up to 140 minutes.  Finally, we have detected enhancements above
the total solar irradiance at the top of the atmosphere of up to
$400\,\text{W}/\text{m}^{2}$, with the $75\,\%$ of the cases below
$200\,\text{W}/\text{m}^{2}$. The most active season of these extreme events is
spring until early summer and autumn. 
-->

@@

For the detection of CE cases we established a baseline of irradiance above which we
characterized each data point as an enhancement event and calculated the over
irradiance (OI). The OI is defined as the irradiance difference of the measured
one-minute GHI from the $\text{GHI}_\text{ref}$ corresponding to cloud-free
atmosphere.
<!-- CE identification criterion ($\text{OI}_i =
\text{GHI}_i - \text{GHI}_\text{CSlim}$), as defined in
Equation\nobreakspace\ref{eq:CE4}. -->
First, we tested two simple approaches for the determination of
$\text{GHI}_\text{ref}$: The Haurwitz's model [@Haurwitz1945], which is a simple
clear sky model and was already adjusted and applied to our data [@Natsis2023].
<!--
, and the total solar irradiance (TSI) at the top of the atmosphere, adjusted for the
Sun-Earth distance.
-->
We have tested both methods by using an appropriate relative
threshold and/or an additional constant offset.  The initial results showed that we
can detect a big portion of the actual CE events.  However, by inspecting the daily
plots of irradiance it became evident that changes in atmospheric conditions
introduced numerous false positive or false negative results. The main reason for
these discrepancies is the variability of the effects of aerosols and water vapor
which were not taken into account in the two simple methods.  To produce a more
representative reference we included the effects of these factors using a radiative
transfer model (RTM).  The applied methodology is discussed in
section\nobreakspace{}\ref{rtmcs}.

>>>>

<!--
We studied the enhancement of the global horizontal irradiance induced by clouds over
Thessaloniki for the period 1993 -- 2023 using data recorded every one minute.  We
identified the cloud enhancement events (CE) by creating an appropriate clear sky
irradiance reference with the use of a radiative transfer model and aerosol optical
depth data from a collocated Cimel sun photometer and a Brewer spectrophotometer. We
found a trend in CE events of
$`r round(yearlyCE[var == "GLB_diff.N", slope], 1)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.N", slope.sd], 1)`\,\text{cases}/\text{year}$,
and a trend in the CE events irradiation of 
$`r round(yearlyCE[var == "GLB_diff.sum", slope], 1)`\pm
`r round(2 * yearlyCE[var == "GLB_diff.sum", slope.sd], 1)`\,\text{kJ}/\text{year}$.
The peak of the CE events was observed during May and June.  An analysis of the total
duration of CE events, showed that durations longer than 5 minutes are very rare,
with exceptions lasting over an hour and up to 140 minutes.  Finally, we have
detected enhancements above the total solar irradiance at the top of the atmosphere
for the same solar zenith angle of up to
$400\,\text{W}/\text{m}^{2}$, with the $75\,\%$ of the cases below
$200\,\text{W}/\text{m}^{2}$. The most active season of these extreme events is
spring until early summer and autumn. 
-->

@@

For the detection of CE cases we established a baseline of irradiance above which we
characterized each data point as an enhancement event and calculated the over
irradiance (OI). The OI is defined as the irradiance difference of the measured
one-minute GHI from the $\text{GHI}_\text{ref}$ corresponding to cloud-free
atmosphere.
<!-- CE identification criterion ($\text{OI}_i =
\text{GHI}_i - \text{GHI}_\text{CSlim}$), as defined in
Equation\nobreakspace\ref{eq:CE4}. -->
First, we used a simple approach for the determination of
$\text{GHI}_\text{ref}$: The Haurwitz's model [@Haurwitz1945], which is a simple
clear sky model and was already adjusted and applied to our data [@Natsis2023].
<!--
, and the total solar irradiance (TSI) at the top of the atmosphere, adjusted for the
Sun-Earth distance.
-->
We created a threshold by using an appropriate relative and/or an additional constant
offset.  The initial results showed that we can detect a big portion of the actual CE
events.  However, by inspecting the daily plots of irradiance it became evident that
changes in atmospheric conditions introduced numerous false positive or false
negative results. The main reason for these discrepancies is the variability of the
effects of aerosols and water vapor which were not taken into account in the two
simple methods.  To produce a more representative reference we included the effects
of these factors using a radiative transfer model (RTM).  The applied methodology is
discussed in section\nobreakspace{}\ref{rtmcs}.

### 2024-08-25 article/article.tex
Commit 2024-08-25 00:10
Commit 2024-08-25 17:12

For the detection of CE cases we established a baseline of irradiance
above which we characterized each data point as an enhancement event and
calculated the over irradiance (OI). The OI is defined as the irradiance
difference of the measured one-minute GHI from the
\(\text{GHI}_\text{ref}\) corresponding to cloud-free atmosphere. First,
we tested two simple approaches for the determination of
\(\text{GHI}_\text{ref}\): The Haurwitz's model \citep{Haurwitz1945},
which is a simple clear sky model and was already adjusted and applied
to our data \citep{Natsis2023}. We have tested both methods by using an
appropriate relative threshold and/or an additional constant offset. The
initial results showed that we can detect a big portion of the actual CE
events. However, by inspecting the daily plots of irradiance it became
evident that changes in atmospheric conditions introduced numerous false
positive or false negative results. The main reason for these
discrepancies is the variability of the effects of aerosols and water
vapor which were not taken into account in the two simple methods. To
produce a more representative reference we included the effects of these
factors using a radiative transfer model (RTM). The applied methodology
is discussed in section\nobreakspace{}\ref{rtmcs}.

@@

{\centering \includegraphics[width=1\linewidth]{../images/P-relative-distribution-diff-1}

@@

\caption{Relative frequency distribution of CE events in Thessaloniki for the period 1993 -- 2023 as a function of OI.}\label{fig:ovir-distribution}
\end{figure}

@@

{\centering \includegraphics[width=1\linewidth]{../images/groups-8}

>>>>

For the detection of CE cases we established a baseline of irradiance
above which we characterized each data point as an enhancement event and
calculated the over irradiance (OI). The OI is defined as the irradiance
difference of the measured one-minute GHI from the
\(\text{GHI}_\text{ref}\) corresponding to cloud-free atmosphere. First,
we used a simple approach for the determination of
\(\text{GHI}_\text{ref}\): The Haurwitz's model \citep{Haurwitz1945},
which is a simple clear sky model and was already adjusted and applied
to our data \citep{Natsis2023}. We created a threshold by using an
appropriate relative and/or an additional constant offset. The initial
results showed that we can detect a big portion of the actual CE events.
However, by inspecting the daily plots of irradiance it became evident
that changes in atmospheric conditions introduced numerous false
positive or false negative results. The main reason for these
discrepancies is the variability of the effects of aerosols and water
vapor which were not taken into account in the two simple methods. To
produce a more representative reference we included the effects of these
factors using a radiative transfer model (RTM). The applied methodology
is discussed in section\nobreakspace{}\ref{rtmcs}.

@@

{\centering \includegraphics[width=1\linewidth]{../images/P-relative-distribution-diff-2}

@@

\caption{Relative frequency distribution of CE events in Thessaloniki for the period 1993 -- 2023 as a function of OI. The histogram was split in two plots with different y-axis scales for better readability.}\label{fig:ovir-distribution}
\end{figure}

@@

{\centering \includegraphics[width=1\linewidth]{../images/groups-7}
